
\documentclass{article}
\usepackage{amssymb}
\usepackage{graphicx}
\usepackage{caption}
\usepackage{subcaption}
\usepackage{listings}
\usepackage{float} %figure inside minipage
\graphicspath{ {./images/} }
\usepackage[export]{adjustbox}
\usepackage{cite}




\begin{document}
\begin{titlepage}
\author{Humaid Agowun and Paul Muir} 
\title{Performance Analysis of ODE Solvers on Covid-19 Models with Discontinuities 
} 
\date{\today} 
\maketitle
\end{titlepage}

\begin{center}
    \textbf{Abstract}
\end{center}

In this report, we consider the numerical solution of two challenging
Covid-19 ordinary differential equation (ODE) 
models that have discontinuities. 
The discontinuities are associated with modeling
the introduction of measures to slow the spread of the virus.
One of the models has a 
time-dependent discontinuity; this means that at a given point in time, a discontinuity is introduced
into the model. The second type of discontinuity is a state-dependent discontinuity;
in this case, the time at which the discontinuity arises depends on the
value of one or more of the solution components, and thus it is not known apriori.
These discontinuities make the models quite challenging for standard ODE solvers to solve.
Furthermore, the presence of exponentially growing solution components adds 
to the difficulties faced by standard ODE solvers when they are used to try to solve these types of 
problems.

In this report, we present an investigation of performance of a collection of ODE solvers (we
consider 21 solvers) available in four popular software environments: R, Python, 
Scilab, and Matlab, when applied to solve these Covid-19 models.

We first focus on straightforward implementations of the models 
where the user employs the solver to attempt to solve the problems using default settings, e.g., default tolerances, and simple implementations for the
discontinuities, i.e., the introduction of `if' statements into the functions that 
define the right-hand sides of the ODE systems. {\it Such implementations of the models
and usage of the solvers are typical of what Covid-19 researchers might employ 
in attempting to solve their models.}

We follow with an investigation of an approach for the solution of the models that uses slightly more
advanced treatments of the models; the approach 
involves making better use of the capabilities of the solvers and 
better implementations of the models themselves.

We also highlight a number of issues with the way that some of the solvers are implemented
in some of the software environments. For example, the treatment of output points, i.e.,  
the points in the domain where solution values should be provided,
is an issue for some of the solvers in some of the software environments. 

{\it We show that the standard use of ODE solvers available within  widely used software 
environments, applied to simple implementations of these Covid-19 models, will
frequently deliver numerical solutions that have no significant digits of accuracy.
Furthermore, the solvers give no indication that the returned solutions
may be inaccurate. We also show that these straightforward treatments of the models 
are always less efficient than the slightly more advanced treatments.
We show that the slightly more advanced treatments of the models can 
result in more efficient computations while at the same time providing
more accurate approximate solutions.}

\include{introduction}
\include{timeProblem}
\include{stateProblem}
\include{exponentialChange}
\include{investigation}
\include{summary}

\bibliographystyle{IEEEtran}
\bibliography{ref}


\end{document}
