\documentclass{article}
\usepackage{amssymb}
\usepackage{graphicx}
\usepackage{caption}
\usepackage{subcaption}
\usepackage{listings}
\usepackage{float} %figure inside minipage
\graphicspath{ {./images/} }
\usepackage[export]{adjustbox}
\usepackage{apacite}




\begin{document}
\begin{titlepage}
\author{Humaid Agowun and Paul Muir} 
\title{Performance Analysis of a Collection of ODE Solvers on Covid-19 Models with Discontinuities 
} 
\date{\today} 
\maketitle
\end{titlepage}

\begin{center}
    \textbf{Abstract}
\end{center}

In this report, we consider the numerical solution of two challenging
Covid-19 ordinary differential equation (ODE) 
models that have discontinuities. One of the models has a 
time-dependent, i.e., at a given point in time, a discontinuity is introduced
into the model. The second type of discontinuity is a state-dependent discontinuity;
this means that the time at which the discontinuity arises depends on the
value of one of the solution components, and thus it is not known apriori.

These discontinuities make the models quite challenging for standard ODE solvers.
Furthermore, the presence of exponentially growing solution components adds further
to the difficulties faced by standard ODE solvers when trying to solve these types of 
problems.

We report on an investigation of performance of a wide array of ODE solvers (we
consider 21 solvers) available in four major software environments: R, Python, 
Scilab, and Matlab, applied to these models.

We focus on straightforward implementations of the models within the above environments
where the user employs the solvers and codes the problems using default settings for 
the solvers, e.g., default tolerances, and simple implementations for the
discontinuities, e.g., the introduction of if-then statements into the functions that 
define the right-hand sides of the ODE systems. Such implementations of the models
and usage of the solvers are typical of what researchers might employ.

We also include an investigation of the solution of the models using slightly more
sophisticated, but easily implementable, treatments of the models; these treatments
involve making better use of the capabilities of the solvers, and slightly more careful
implementations of the models themselves.

We also highlight a number of issues with the way that some of the solvers are implemented
in some of the software environments. For example, the way in which output points, where 
the user specifies points in the domain where the solution value should be provided,
is an issue in some of the software environments. 

{\it We show that the standard use of ODE solvers available within  widely used software 
environments, applied to simple implementations of these Covid-19 models, will
frequently deliver numerical solutions that have no significant digits of accuracy.
Furthermore, the solvers give no indication that the return solutions
may be inaccurate.} We also show that the straightforward treatments of the models 
are always less efficient than the slightly more sophisticated treatments.
{\it We show that the slightly more sophisticated treatments of the models can 
result in more efficient computations while at the same time providing
much more accurate solutions.}

\include{introduction}
\include{timeProblem}
\include{stateProblem}
\include{investigation}
\include{summary}

\bibliographystyle{ieeetran}
\bibliography{bibliography}
\appendix
\include{ebolaPaper}

\end{document}
